```{r echo=FALSE}
opts_chunk$set(fig.path = 'figure/fruit/', 
               fig.height = 6, 
               fig.width = 12, 
               tidy = FALSE, 
               warning = FALSE,
               cache = TRUE)
read_chunk('fruit_biomass.R')
```

# Phenolgy and Biomass

******

<br>

Source files available on [github](https://github.com/camposfa/Transects)

# Major steps
1. Load phenology data and clean it up: [jump](#id1)
2. Smooth phenology data to obtain monthly pheno indices: [jump](#id2)
3. Load transect data and clean it up: [jump](#id3)
4. Specify minimum DBH thresholds for phenology species in transects: [jump](#id4)
5. Estimate maximum potential biomass for each species: [jump](#id5)
6. Combine phenology with transect data to obtain actual monthly biomass estimates: [jump](#id6)

Prepare R environment
```{r workspace, message = FALSE, results = 'hide'}
```

******

<br>

<a id="id1"></a>
## Step 1: Load phenology data and clean it up

### Obtain Data

#### Option 1
- If you have access, pull current data from PACE database
- Must create SSH tunnel first using plink
- Change dsn="PACE" to whatever your PACE ODBC connection is called
- _Remove comments if using!_
```{r pheno_read_pace}
```

#### Option 2
- [Contact me](mailto:camposfa@gmail.com) if you are interested in the raw data
```{r pheno_read_csv}
```


### Clean up data
```{r pheno_clean}
```


### Unique trees for each pheno species
```{r pheno_table_count_species, cache = TRUE, echo = FALSE , results = 'asis'}
```

******

### Calculate pheno indices for each phenology record
- index_avail = (Maturity / 4) x (Coverage / 4)
- index_avail is maximum (1) if full coverage and maturity
- index_avail is zero if either coverage or maturity is zero
```{r pheno_index}
```

### Filter pheno data (remove some species)
```{r pheno_filter}
```

### Monthly plot of index_availability indices for selected species
```{r pheno_plot_index, fig.height = 12}
```

******

<br>

<a id="id2"></a>
## Step 2: Smooth phenology data to obtain monthly pheno indices

### GAM smoothing of phenology data using availability index
- Applied to each species separately
- Smoothed with cyclic cubic regression spline, which forces the ends (January / January) to match
- Using tree_id as random effect
```{r pheno_smooth}
```

### Plot all smoothed pheno index_availability estimates
```{r pheno_plot_smooth, fig.height = 12}
```

Pheno data for some species looks very problematic, notably ARET, RMON, SOBO, TAME, and TOCH

******

<br>

<a id="id3"></a>
## Step 3: Load transect data and clean it up

### Load transect data
- [Contact me](mailto:camposfa@gmail.com) if you are interested in the raw data
```{r tr_read_csv}
```

### Clean up transect data
```{r tr_clean}
```

******

<br>

<a id="id4"></a>
## Step 4: Specify minimum DBH thresholds for phenology species in transects

### Load FPV data
- [Contact me](mailto:camposfa@gmail.com) if you are interested in the raw data
```{r fpv_read_csv}
```

### Clean up the FPV data
```{r fpv_clean}
```

### Find minimum DBHs from FPV data
```{r fpv_min_dbh}
```

### Plot DBHs for each species to identify problem cases
```{r fpv_plot_dbh}
```

### Set min dbh for problem species case-by-case
- I tried automating this, but it didn't work well
- If you want to see the detail, you can contact me or  check the "fpv_fix_dbh_verbose" chunk of code in the source file
```{r fpv_fix_dbh_brief}
```

### Adjust minimum dbh thresholds for 3 additional species: TAME, SEXC, SGLN
These thresholds have been fudged slightly to include the largest tree(s) of that species in the transects, each of which falls just below the FPV-based cutoff. 
```{r fpv_dbh_extra}
```

### Subset tr_pheno to include only pheno species
```{r tr_filter}
```

### Determine if each tree in the transect data meets the minimum DBH criteria
```{r tr_test_usable}
```

### Number of usable trees in transects
```{r tr_table_usable, cache = TRUE, echo = FALSE , results = 'asis'}
```

******

<br>

<a id="id5"></a>
## Step 5: Estimate maximum potential biomass for each species

### Biomass estimates
- Assumes every tree is fruiting at max potential
- Calculated by applying Peters 1998 formula on each tree and multiplying by the proportion of the tree in transect, then summed for each species
```{r biomass_total}
```

# Plot max potential biomass for each species
```{r biomass_plot_max}
```

******

<br>

<a id="id6"></a>
## Step 6: Combine phenology with transect data to obtain actual monthly biomass estimates

### Merge fruit index_availability and and biomass data sets
```{r biomass_available1}
```

### Aggregate to obtain sum for all species in each month
```{r biomass_monthly1}
```

### Plot of monthly biomass estimates
```{r biomass_plot_monthly1}
```

The data are strongly affected by a few large, common species that are either very rarely eaten or have very small, wind-dispersed seeds for which Peters' biomass formula is inappropriate. Therefore, I'm repeating the calculation without these species.

### Without GULM and wind-dispersed seeds: LCAN, LSPE, BUNG, CCAN
```{r biomass_available2}
```

### Plot of monthly biomass estimates
```{r biomass_plot_monthly2}
```

### Total fruit production per hectare over entire year (excluding GULM and LCAN)
```{r biomass_sum}
```

### Interpolation of monthly to daily biomass
```{r bioimass_interpolation}
```

### Plot interpolated biomass over entire annual cycle
```{r biomass_interpolation_plot}
```

### Calculate over my 2010 - 2011 study period
```{r biomass_study_period}
```

### Plot daily biomass estimate over my study period
```{r biomass_study_period_plot}
```